### 2022-04-30 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@e8f7f88e898980f4260ab03e9f584f170625e1fa 🚀

\hypertarget{host-distribution-does-not-match-virus-distinctiveness}{%
\subsection{Host distribution does not match virus
distinctiveness}\label{host-distribution-does-not-match-virus-distinctiveness}}

Bats are found worldwide and are one of the most diverse groups among
mammals (Moratelli \& Calisher, 2015), and one of the main animal
reservoir for different strains of betacoronaviruses (Drexler et al.
2014). This has attracted attention to areas where high diversity of
bats, and therefore presumably high diversity of betacoronaviruses, can
be an important issue for human health (Calisher et al. 2006, Moratelli
and Calisher 2015). Accordingly, we collected the IUCN rangemaps for
known hosts of betacoronaviruses, to illustrate where hotspots of host
diversity are. These results are presented in fig.~\ref{fig:richness}
(top). As per our current knowledge of which bats are hosts of
betacoronaviruses, these hotspots are primarily South-East Asia, parts
of Europe, and to a lesser extent parts of Africa in the -25-0 range of
latitudes. Previous research (\textbf{Anthony et al., 2017; Mollentze \&
Streicker, 2020}) states that locally diverse bat communities could
maintain more viruses and hence, a higher probability of having a
pathogen that could represent a risk for human health; locally diverse,
virus-rich bats communities could represent an increased risk of
spillover under climate change (\textbf{Ice ice berg berg}). This
probability involves multiple factors, among which the relatedness of
hosts (which can make the jumps easier (\textbf{Longdon et al., 2011;
Mollentze et al., 2020; Wolfe et al., 2007}), and the overall tendency
of hosts within a locality to share viruses, which may limit viral
diversity because of within-host competition (Leeks et al., 2018;
Sallinen et al., 2020). All things considered, the richness of known
betacoronaviruses hosts is not a sufficient predictor of spillover risk,
and we should move towards a more coevolutionary understanding of this
system.

\begin{figure}
\hypertarget{fig:richness}{%
\centering
\includegraphics{figures/combined_richness.png}
\caption{Top panel: relative diversoty of known bat hosts of
betacoronaviruses. This map shows that the region with the largest
number of possible hosts is South-Eastern Asia. Bottom panel: congruence
between the evolutionary distinctiveness of the hosts (grey to blue) and
the viruses (grey to red). By contrast to the richness map, this reveals
that South America has the most evolutionary distinct hosts \emph{and}
viruses, whereas South-Eastern Asia has mostly distinct viruses. This is
congruent with know results about New World bats being evolutionary
distinct, and suggests that they similarly have distinct
viruses.}\label{fig:richness}
}
\end{figure}

High density of hosts sharing the same virus (albeit possibly different
strains) calls into question the evolution of the bat antiviral immune
system and its co-evolution with viruses, which may result in distinct
immunological responses in different areas, as evidenced in other bat
species (Banerjee et al. 2020). Immune characteristics that allow bats
to be better adapted to infection by emerging viruses (Gorbunova et al.,
2020; Irving et al., 2021) may be related to a wide variety of diets
(Jones et al., 2022; Moreno Santillán et al., 2021; Banerjee et al.,
2020; Schneeberger et al., 2013). Considering whether viruses easily
adapted to multiple hosts have lower virulence on these hosts, or lower
ability to jump to hosts with different immune characteristics, should
yield valuable additional predictors for the total risk of spillover.
The subset of chiroptera that are hosts of betacoronaviruses fits the
evolutionary timeline of the group. Chiropterans can be classified as
Microchiroptera and macrochiroptera, where macrochiroptera have an older
history from an evolutionary perspective compared to macrochiroptera
(Teeling et al. 2005, Springer 2013). Specifically, we would expect that
the New World group of bats, being more evolutionary distinct, would
also have evolutionary distinct viruses. Indeed fig.~\ref{fig:richness}
(bottom) reveals it to be the case. By contrast, South-Eastern Asia has
a lot of non-evolutionary distinct bats, but evolutionary-distinct
viruses.

Viral evolutionary distinctiveness largely tracks host diversity,
particularly in southern China but, oddly, not throughout the rest of
southeast Asia. This indicates, perhaps, that many distinctive viruses
remain to be discovered in this region (an idea that is unsurprising
given the growing realization, around the emergence of SARS-CoV-2, that
a unique lineage of similar viruses are widespread in bats but still
mostly undescribed). The most distinct betacoronaviruses are found in
South America, a region with a comparatively lower number of hosts; this
suggests that the South American bat-betacoronvirus complex has been
more isolated, and is probably undergoing a different co-evolutionary
dynamic. Alternatively, this distinctiveness hostpot may be a product of
under-sampling: South-America is one of the places where the fewest
betacoronaviruses have been discovered (Anthony et al. 2017), and adding
more viruses would bring the distinctiveness of known sequences down.
Previous work has suggested the Americas may be a hotspot of both
undiscovered bat viruses in general (Olival et al. 2017, Allen et al.
2017) and coronavirus specifically (Anthony et al. 2017), though not
necessarily betacoronaviruses, and particularly not those in clades with
notable zoonotic potential.

@@

@@

@@

In order to turn the hypotheses based on the Geographic Mosaic Theory of
Coevolution into a measure of risk, we overlapped three components of
spillover risk: viral sharing, \emph{i.e.} the chance that two bats will
share viruses overall; Local Contribution to Beta Diversity, \emph{i.e.}
the fact that a bat community is compositionally unique compared to the
average compositional similarity across the entire system; finally, the
phylogenetic diversity, \emph{i.e.} how dispersed the bats in a location
are within the tree of life. These results are presented using an
additive color mapping in Figure xx, and lead to the definition of broad
biogeographic regions of risk, where the same color represents the same
type of risk. Pairwise maps of the three components are present in
supplementary materials.

@@

Finally, we provide a summary visualization of what available
information describes the spillover of zoonotic betacoronaviruses of bat
origin where data was available before and up through the COVID-19
pandemic. The SARS-CoV-2 outbreak was georeferenced to the initial case
cluster in Wuhan, China; SARS-CoV was georeferenced based on the cave
with the closest known viruses circulating in nature (Hu et al.~2017
PLoS Pathogens), and a nearby location where serological (antibody)
evidence has indicated human exposure to SARS-like viruses (Wang et
al.~2018 Virologica Sinica). For MERS-CoV, we presented the index cases
available from a recently-published compendium of MERS-CoV cases
(Ramshaw et al.~2019); these are largely if not all presumed to be
camel-to-human transmission, and the precise origin point of MERS-CoV in
bats is uncertain. Not shown is a recent case of a recombinant canine
coronavirus that showed the ability to infect humans, both because this
study was published after the beginning of the COVID-19 pandemic and
because bats' involvement in this cycle of transmission has been
marginal to non-existent.

@@

>>>>

\hypertarget{host-richness-does-not-predict-virus-distinctiveness}{%
\subsection{Host richness does not predict virus
distinctiveness}\label{host-richness-does-not-predict-virus-distinctiveness}}

Bats are found worldwide and are both one of the most diverse groups
among mammals (\textbf{Moratelli \& Calisher, 2015}), and one of the
main animal reservoir for different strains of betacoronaviruses
(Drexler et al. 2014). This has attracted attention to areas where high
diversity of bats, and therefore presumably high diversity of
betacoronaviruses, can be an important issue for human health (Calisher
et al. 2006, Moratelli and Calisher 2015). By overlaying the IUCN
rangempas for confirmed bat hosts of betacoronaviruses
{[}fig.~\ref{fig:richness}; top{]}, we see that the the main hotspots of
host richness are primarily South-Eastern Asia, parts of Southern
Europe, and to a lesser extent parts of Africa in the -25-0 range of
latitudes. The description of host richness is an important first step
towards understanding risk, as previous research (\textbf{Anthony et
al., 2017; Mollentze \& Streicker, 2020}) states that locally diverse
bat communities could maintain more viruses and hence, a higher
probability of having a pathogen that could represent a risk for human
health.

\begin{figure}
\hypertarget{fig:richness}{%
\centering
\includegraphics{figures/combined_richness.png}
\caption{Top panel: relative diversity of known bat hosts of
betacoronaviruses. This map shows that the region with the largest
number of possible hosts is South-Eastern Asia. Bottom panel: congruence
between the evolutionary distinctiveness of the hosts (grey to blue) and
the viruses (grey to red). By contrast to the richness map, this reveals
that South America has the most evolutionary distinct hosts \emph{and}
viruses, whereas South-Eastern Asia has mostly distinct viruses. This is
congruent with know results about New World bats being evolutionary
distinct, and suggests that they similarly have distinct
viruses.}\label{fig:richness}
}
\end{figure}

Nevertheless, locally diverse and virus-rich bat communities could
represent an increased risk of spillover under climate change through
the creation of novel interactions (\textbf{Ice ice berg berg}), and
therefore the diversity of betacoronavirus strains should similarly be
ccounted for. In fig.~\ref{fig:richness} (bottom), we contrast the
evolutionary distinctiveness of bats and viruses -- this reveals a
slightly different portrait than bat richness alone. Chiropterans can be
classified, from a macro-evolutionary standpoint, as microchiroptera and
macrochiroptera, where macrochiroptera have an older history from an
evolutionary perspective compared to macrochiroptera (Teeling et al.
2005, Springer 2013). Specifically, we would expect that the so-called
``New World'' group of bats, being more evolutionary distinct, would
also have evolutionary distinct viruses. Indeed fig.~\ref{fig:richness}
(bottom) reveals it to be the case, and this region harbors a distinct
bat-betacoronavirus complex. By contrast, South-Eastern Asia has a lot
of non-evolutionary distinct bats, but evolutionary-distinct viruses.

It is noteworthy that outside of South America, viral evolutionary
distinctiveness does not accurately tracks host diversity, with some
areas having over-distinct viruses (southern China but, oddly, not the
rest of southeast Asia). There are a number of likely explanations.
First, given the richness of bats in southeast Asia, many
betacoronaviruses likely remain to be discovered in this region. Indeed,
global predictions by (\textbf{Becker?}) highlight that southeast Asia
is a likely hostpot of unconfirmed hosts of betacoronaviruses, which
would likely result in additiona viral discoveries. This idea is
unsurprising given the growing realization, around the emergence of
SARS-CoV-2, that a unique lineage of similar viruses are widespread in
bats but still mostly undescribed. The most distinct
bats/betacoronavirus complex is found in South America, a region with a
comparatively lower number of hosts; this matches with the isolation
through viariance of the host group, and may highlight a different
co-evolutionary dynamic. Alternatively, this distinctiveness hostpot may
be a product of under-sampling: South-America is one of the places where
the fewest betacoronaviruses have been discovered (Anthony et al. 2017,
Olival et al. 2017, Allen et al. 2017), resulting in sparser
phylogenetic tree, thereby artificially inflating distinctiveness.
Adding more viruses would bring the distinctiveness of known sequences
down.

@@

Despite the difference in evolutionary distinctiveness globally, there
are reasons to expect that the phylogeography of bats and
betacoronaviruses should show some degree of congruence. High density of
hosts sharing the same virus (albeit possibly different strains) can
drive or result from evolution of the bat antiviral immune system,
resulting in spatially distinct immunological responses, as evidenced in
several bat species (Banerjee et al. 2020). Immune characteristics that
allow bats to be better adapted to infection by emerging viruses
(Gorbunova et al., 2020; Irving et al., 2021) may be related to a wide
variety of diets (Jones et al., 2022; Moreno Santillán et al., 2021;
Banerjee et al., 2020; Schneeberger et al., 2013), themselves likely to
be driven by spatial effect, especially at the local scale -- bats,
indeed, occupy a variety of environments.

@@

In fig.~\ref{fig:biogeo}, we show a projection of the phylogeographic
signal of bats (top) and viruses (bottom) in space; the distinct
groupings (represented by different colors symbolizing positions in the
subspace formed by the first two axes of the PCoA) are essentially
equivalent between the two groups, and can be coarsely delineated as
southeast Asia, Eurasia above a northing of 25, and Africa and south
America. These results suggest that, although the evolutionary
distinctiveness of the bat/betacoronavirus complex varies spatially, the
system shows an important degree of spatial consistency, with a reduced
number of bioregions. Available information describing the spillover of
zoonotic betacoronaviruses of bat origin where data was available before
and up through the COVID-19 pandemic puts spillover events of SARS-CoV-2
in Wuhan, China; SARS-CoV in \textbf{XXX} based on the presence of
closest known viruses circulating in nature, and a nearby location where
serological (antibody) evidence has indicated human exposure to
SARS-like viruses (Wang et al.~2018 Virologica Sinica); MERS-CoV in
\textbf{XXX} based on index cases available from a recently-published
compendium of cases (Ramshaw et al.~2019). For the latest event, most if
not all index cases are presumed to be camel-to-human transmission, and
the precise origin point (if it exists) of MERS-CoV in bats is
uncertain. Recent recombinant canine coronavirus spillover events in
Haiti (\textbf{ref}) and Europe (\textbf{ref}) are not relevant here, as
bats' involvement in these cycles of transmission have been supposed to
be non-existent.

@@

As host richness, joint disctinveness, or phylogeographic structure
suggest that the bat/betacoronavirus complex is globally fragmented
enough to give rise to both different levels of risk (as evidenced by
the spatial location of spillover events) and different types of
co-evolutionary dynamics, we turn to the Geographic Mosaic Theory of
Coevolution \textbf{REF} to provide a measure of risk accounting for
multiple processes. In fig.~\ref{fig:trivariate}, we overlapped three
components of spillover risk: viral sharing, \emph{i.e.} the chance that
two bats will share viruses overall; Local Contribution to Beta
Diversity, \emph{i.e.} the fact that a bat community is compositionally
unique compared to the average compositional similarity across the
entire system; finally, host phylogenetic diversity, \emph{i.e.} how
dispersed the bats in a location are within the tree of life. This
approach leads to the definition of broad biogeographic regions of risk,
where the same color represents the same type of risk. By way of
constrat to figures fig.~\ref{fig:richness} and fig.~\ref{fig:biogeo},
these regions do not necessarilly overlap with previous spatial
partitions of the bat/betacoronavirus complex.

@@

@@

This probability involves multiple factors, among which the relatedness
of hosts (which can make the jumps easier (\textbf{Longdon et al., 2011;
Mollentze et al., 2020; Wolfe et al., 2007}), and the overall tendency
of hosts within a locality to share viruses, which may limit viral
diversity because of within-host competition (Leeks et al., 2018;
Sallinen et al., 2020). All things considered, the richness of known
betacoronaviruses hosts is not a sufficient predictor of spillover risk,
and we should move towards a more coevolutionary understanding of this
system.

Considering whether viruses easily adapted to multiple hosts have lower
virulence on these hosts, or lower ability to jump to hosts with
different immune characteristics, should yield valuable additional
predictors for the total risk of spillover.
